---
title: "feature_selection_stability"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{feature_selection_stability}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This document compares the stability of feature selection based on specific performance measures. 

First, we create a custom feature selection method that performs sequential forward search based on an arbitrary performance measure using `mlr` library.

```{r}
custom.featsel <- function(x, y, q, measure){
  library(mlr)
  dataset <- cbind.data.frame(x, y)
  dataset.task <- makeClassifTask(data=dataset, target = 'y')
  rdesc <- makeResampleDesc("Holdout")
  lrn <- makeFeatSelWrapper(makeLearner("classif.plr", predict.type='prob'),
                           resampling = rdesc, 
                           measures = measure,
                           control = makeFeatSelControlSequential(method='sfs', alpha=0), 
                           show.info = FALSE)  
  capture.output(
    mod <- train(lrn, task=dataset.task), 
    file='NUL')
  
  fsel_res <- getFeatSelResult(mod)
  selection_path <- fsel_res$x
  all_features <- mod$features
  features_not_selected <- all_features[!(all_features %in% selection_path)]
  selection_path_extended <- c(selection_path, sample(features_not_selected))
  
  path_matrix <- rep(F, length(selection_path_extended))
  for (i in 1:length(selection_path_extended)){
    path_matrix <- rbind(path_matrix,
                         all_features %in% selection_path_extended[1:i])
  }
  colnames(path_matrix) <- all_features
  path_matrix <- t(path_matrix)
  colnames(path_matrix) <- sapply(1:ncol(path_matrix), function(x){paste0('s',x-1)})
  path_matrix
  selected_features <- all_features %in% selection_path
  names(selected_features) <- all_features
  return(list(selected = selected_features, path = path_matrix))
}
```


```{r}
plot_stabsels <- function(phat, title, red_id, black_id){
  to_plot <- phat
  plot(to_plot[1,], type='l', col='red',
       ylab="Selection probability", 
       xlab="N variables selected",
       main=title)
  for(i in red_id){
    lines(to_plot[i,], type='l', col='red')
  }
    for(i in black_id){
    lines(to_plot[i,], type='l', col='black')
  }
}
```

Next, run the stability selection procedure from `stabs` library using our custom feature selection method for specific performance measures

```{r}
custom.featsel.acc <- function(x, y, q){
  custom.featsel(x, y, q, mlr::acc)
}
custom.featsel.auc <- function(x, y, q){
  custom.featsel(x, y, q, mlr::auc)
}
custom.featsel.brier <- function(x, y, q){
  custom.featsel(x, y, q, mlr::brier)
}
custom.featsel.logloss <- function(x, y, q){
  custom.featsel(x, y, q, mlr::logloss)
}

# source('./functions.R')
downsample_to_balance <- function(y){
     nr <- length(y)
     size <- min(table(y))
     idx <- lapply(split(seq_len(nr), y), function(.x) sample(.x, size))
     unlist(idx)
}
```


```{r}
data("PimaIndiansDiabetes")
ds <- PimaIndiansDiabetes
ds$diabetes <- as.factor(as.numeric(ds$diabetes) - 1)
downsampled_idx <- downsample_to_balance(ds$diabetes)
ds <- ds[downsampled_idx,]


x <- ds[,1:8]
x_permuted <- x
x_permuted <- apply(x_permuted, 2, function(x){sample(x)})
colnames(x_permuted) <- sapply(colnames(x_permuted), function(x){paste0('perm_', x)})
x <- cbind.data.frame(x, x_permuted)
y <- ds[,9]


library(stabs)
custom.stabsel.acc.pima <- stabsel(x,
                              y,
                              cutoff=0.75,
                              PFER=1,
                              fitfun=custom.featsel.acc)
custom.stabsel.auc.pima <- stabsel(x,
                              y,
                              cutoff=0.75,
                              PFER=1,
                              fitfun=custom.featsel.auc)
custom.stabsel.brier.pima <- stabsel(x,
                              y,
                              cutoff=0.75,
                              PFER=1,
                              fitfun=custom.featsel.brier)
custom.stabsel.logloss.pima <- stabsel(x,
                              y,
                              cutoff=0.75,
                              PFER=1,
                              fitfun=custom.featsel.logloss)
```


```{r}
par(mfrow=c(2,2))
plot_stabsels(custom.stabsel.acc.pima$phat, 'Accuracy', 1:8, 9:16)
plot_stabsels(custom.stabsel.auc.pima$phat, 'auc', 1:8, 9:16)
plot_stabsels(custom.stabsel.brier.pima$phat, 'brier', 1:8, 9:16)
plot_stabsels(custom.stabsel.logloss.pima$phat, 'Logarithmic', 1:8, 9:16)
```


```{r}

library(kernlab)
data("spam")
spam$type <- as.numeric(spam$type) - 1
names(spam)[58] <- 'y'

spam <- spam[1:2000,]
spam <- spam[downsample_to_balance(spam$y),]

x <- spam[,1:20]
x_permuted <- x
x_permuted <- apply(x_permuted, 2, function(x){sample(x)})
colnames(x_permuted) <- sapply(colnames(x_permuted), function(x){paste0('perm_', x)})
x <- cbind.data.frame(x, x_permuted)
y <- as.factor(spam$y)


custom.stabsel.acc.spam <- stabsel(x,
                              y,
                              cutoff=0.75,
                              PFER=1,
                              fitfun=custom.featsel.acc)
custom.stabsel.auc.spam <- stabsel(x,
                              y,
                              cutoff=0.75,
                              PFER=1,
                              fitfun=custom.featsel.auc)
custom.stabsel.brier.spam <- stabsel(x,
                              y,
                              cutoff=0.75,
                              PFER=1,
                              fitfun=custom.featsel.brier)
custom.stabsel.logloss.spam <- stabsel(x,
                              y,
                              cutoff=0.75,
                              PFER=1,
                              fitfun=custom.featsel.logloss)
```


```{r}
par(mfrow=c(2,2))
plot_stabsels(custom.stabsel.acc.spam$phat, 'Accuracy', 1:20, 21:40)
plot_stabsels(custom.stabsel.auc.spam$phat, 'auc', 1:20, 21:40)
plot_stabsels(custom.stabsel.brier.spam$phat, 'brier', 1:20, 21:40)
plot_stabsels(custom.stabsel.logloss.spam$phat, 'Logarithmic', 1:20, 21:40)
```

plot

```{r}
library(reshape2)
library(ggplot2)
library(cowplot)

format_stabsel_data <- function(stabsel_data, measure){
  data_to_plot <- as.data.frame(t(stabsel_data$phat))
  data_to_plot$n_selected <- as.numeric(unlist(lapply(row.names(data_to_plot), function(x) {strsplit(x, 's')[[1]][2]})))
  data_to_plot <- melt(data_to_plot, id.vars = 'n_selected')
  data_to_plot$measure = measure
  return(data_to_plot)
}

pubtheme <- theme(aspect.ratio = 1,
                  text=element_text(size=7),
                  plot.title = element_text(size=8),
                  axis.text = element_text(size=7), 
                  strip.text = element_text(size=8, face='bold'),
                  # legend.position = 'none',
                  legend.justification = c(0.5, 0.5))


data_to_plot_spam_long <- do.call(rbind, 
                             mapply(format_stabsel_data, 
                             list(custom.stabsel.acc.spam,
                                  custom.stabsel.auc.spam,
                                  custom.stabsel.brier.spam,
                                  custom.stabsel.logloss.spam),
                             list('Accuracy', 'AUC', 'Brier', 'Logscore'),
                             SIMPLIFY = FALSE))
data_to_plot_spam_long$permuted <- ifelse(
  startsWith(as.character(data_to_plot_spam_long$variable), 'perm'),
  'Permuted', 'Not permuted')


p_spam <- ggplot(data_to_plot_spam_long, 
       aes(x=n_selected, y=value, group=variable, color=permuted)) +
  geom_line() +
  facet_wrap(~measure, nrow = 1) +
  scale_color_manual(values = c('red', 'black')) +
  ylab('Selection probability') +
  xlab('Number of variables selected') +
  theme_minimal_grid() +
  theme(legend.title = element_blank()) +
  pubtheme 
```


```{r}
data_to_plot_pima_long <- do.call(rbind, 
                             mapply(format_stabsel_data, 
                             list(custom.stabsel.acc.pima,
                                  custom.stabsel.auc.pima,
                                  custom.stabsel.brier.pima,
                                  custom.stabsel.logloss.pima),
                             list('Accuracy', 'AUC', 'Brier', 'Logscore'),
                             SIMPLIFY = FALSE))
data_to_plot_pima_long$permuted <- ifelse(
  startsWith(as.character(data_to_plot_pima_long$variable), 'perm'),
  'Permuted', 'Not permuted')

p_pima <- ggplot(data_to_plot_pima_long, 
       aes(x=n_selected, y=value, group=variable, color=permuted)) +
  geom_line() +
  facet_wrap(~measure, nrow = 1) +
  scale_color_manual(values = c('red', 'black')) +
  ylab('Selection probability') +
  xlab('Number of variables selected') +
  theme_minimal_grid() +
  theme(legend.title = element_blank()) +
  pubtheme 
```


```{r}
library(ggpubr)

p_stabsel <- ggarrange(p_spam,
          p_pima,
          nrow = 2, 
          common.legend = TRUE,
          legend = 'bottom',
          # labels = 'AUTO',
          labels = c('Diabetes data', 'Spam data'),
          font.label = list(size=8))
p_stabsel

# ggsave('stabsel.tiff', p_stabsel, width = 6, height = 4, units = 'in')
```


