---
title: "power_against_alternative"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{power_against_alternative}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

In this document, we compare type I and type II error rates when comparing the performance of competing models using various performance measures. This largely mirrors the analysis in `power_against_chance_level.Rmd` but we are not testing the hypothesis that there is no effect, but that the two models have the same performance (i.e., the difference between their errors is 0)

# Set up helper functions

A convenience function to compute multiple p values, the actual functions that compute p values are in the `../helpers/significance_comparisons_functions.R`

```{r setup}
source('../helpers/data_loading.R')
source('../helpers/significance_comparisons_functions.R')
source('../helpers/utility_functions.R')
source('../helpers/scoring_functions.R')

compare_all <- function(preds1, preds2, y){
  c('acc'=compare_accuracy(preds1 < 0.5, preds2 < 0.5, y),
    'auc'=compare_auc(preds1, preds2, y),
    'brier'=compare_brier(preds1, preds2, y),
    'logscore'=compare_logscore(preds1, preds2, y))
}
```

We use the same function as in `power_against_chance_level.Rmd`

```{r, cache=TRUE}
library(parallel)
calculate_proportion_signifficant <- function(generator,
                                              parameters,
                                              n_sims, 
                                              ncores=1){
  # intialize cluster for parallel processing, type='FORK' for workers
  # to be able to use defined functions, i.e. the generator
  cl = makeCluster(ncores, type = 'FORK')
  clusterSetRNGStream(cl) # so that random numebrs don't synchronize 
  on.exit(stopCluster(cl))
  # for each parameter form the list of parameters, we run generator function
  # that produce p-values n_sims number of times
  results <- list()
  for(parameter in parameters){
    # this does the same as the line bellow, only in parallel
    # pvals <- t(replicate(n_sims, generator(param)))
    pvals <- t(parSapply(cl, 1:n_sims, function(x){generator(parameter)}))
    # calculate proportions of p values < 0.05
    results <- append(results, list(colMeans(pvals < 0.05)))
  }
  results <- as.data.frame(do.call(rbind, results))
  results$parameter <- parameters
  return(results)
}
```

and the plotting function

```{r, cache=TRUE}
library(ggplot2)
library(cowplot)
library(reshape2)

plot_results_lines <- function(alphas, id_name='parameter'){
  lines_colors <- c("black", "#375e97", "#fb6542", 'black')
  melted_alphas <- melt(alphas, id.vars = id_name)
  names(melted_alphas)[1] <- 'x'
  
  ggplot(melted_alphas, aes(x=x, y=value, color=variable, linetype=variable)) +
    geom_line()  +
    geom_point(aes(shape=variable, size=variable)) +
    scale_color_manual(values=lines_colors) +
    scale_linetype_manual(values = c('solid', 'solid','solid',  'dashed')) +
    scale_shape_manual(values=c(19,19,19,1)) +
    scale_size_manual(values=c(2,2,2,4)) +
    
    theme_minimal_grid() +
    scale_y_continuous(limits=c(0,1), breaks=seq(0, 1, 0.2))+
    theme(aspect.ratio = 1,
          # text=element_text(size=7),
          # axis.text = element_text(size=7),
          legend.position = 'bottom',
          legend.justification = c(0.5, 0.5),
          legend.title=element_blank(),
          legend.key.width = unit(20, 'mm'))  +
    ylab('Proportion p < 0.05') +
    xlab('Relative training set size')
}
```


### Simulated predictions

Create a function that simulates correlated predictions from two models. We can specify the performance of the first model and how much is the second model better.

```{r, cache=TRUE}
library(psych)

simulate_comparison <- function(n, signal_sd, random_difference_sd,
                                improvement_scale){
  y <- c(rep(0, n), rep(1, n))
  predictions_latent <- c(rnorm(n, -signal_sd/2), rnorm(n, signal_sd/2))
  predictions_noise <- rnorm(length(predictions_latent), sd=random_difference_sd)
  predictions_improvement <- sign(y-0.5)*improvement_scale
  predictions_A <- logistic(predictions_latent + predictions_noise)
  predictions_B <- logistic(predictions_latent - predictions_noise + predictions_improvement)
  compare_all(predictions_A, predictions_B, y)
}
```

and we create a generator out of this

```{r, cache=TRUE}
gen_compr_pval_n <- function(n){
  simulate_comparison(n = n,
                      signal_sd = 1,
                      random_difference_sd = 0.1,
                      improvement_scale = 0.05)
}

gen_compr_pval_n_null <- function(n){
  simulate_comparison(n = n,
                      signal_sd = 1,
                      random_difference_sd = 0.1,
                      improvement_scale = 0) # true improvement == 0
}
```

and run the analysis

```{r, cache=TRUE}
n_sims <- 1000

n_vec <- seq(20,500,100)
results_comp_n <- calculate_proportion_signifficant(gen_compr_pval_n,
                                                    parameters = n_vec,
                                                    n_sims = n_sims,
                                                    ncores = 4)
plot_results_lines(results_comp_n)
```


```{r, cache=TRUE}
results_comp_n_null <- calculate_proportion_signifficant(gen_compr_pval_n_null,
                                                    parameters = n_vec,
                                                    n_sims = n_sims*10,
                                                    ncores = 4)
plot_results_lines(results_comp_n_null)
```

### Simulated data - GPR predictions

A function that takes the data generator, number of subjects, and the proportion of smaller training set to fit two models and compare their predictions.

```{r, cache=TRUE}
compare_gpc_full_vs_reduced <- function(data_generator, n, prop, 
                                        shuffle.predictions=FALSE){
  library(caret)
  library(kernlab)
  data <- data_generator(n*2)
  data <- as.data.frame(data)
  train_idx <- createDataPartition(data$y, list=F)
  test_data <- data[-train_idx,]
  train_data <- data[train_idx,]
  
  train_idx_small <- createDataPartition(train_data$y, list=F, p=prop)
  train_data_small <- train_data[train_idx_small,]
  
  fit_full <- gausspr(as.factor(y) ~ .,
                      data=train_data,
                      kpar=list(sigma=0.01))
  
  fit_reduced <- gausspr(as.factor(y) ~ .,
                      data=train_data_small,
                      kpar=list(sigma=0.01))
  
  predictions_full <- as.data.frame(predict(fit_full,
                                             newdata=test_data,
                                             type='probabilities'))$`1`
  predictions_reduced <- as.data.frame(predict(fit_reduced,
                                             newdata=test_data,
                                             type='probabilities'))$`1`
  if (shuffle.predictions){
     shuffled_idx <- sample(x = c(FALSE, TRUE), 
                            replace = T, 
                            prob = c(0.5, 0.5), 
                            size = length(predictions_full))
     predictions_all <- cbind(predictions_full, predictions_reduced)
     predictions_all[shuffled_idx,] <- t(apply(predictions_all[shuffled_idx,],
                                             1,
                                             rev))
     predictions_full <- predictions_all[,1]
     predictions_reduced <- predictions_all[,2]
  }
  compare_all(predictions_full, predictions_reduced, test_data$y)
}
```

simulated data data generator

```{r, cache=TRUE}
mixed_gaussians_data_generator <- function(n){
  library(MASS)
  n_signal <- 50
  n_noise <- 50
  signal_strenght <- 0.4

  sigma <- diag(n_signal + n_noise)
  x1 <- mvrnorm(n=n, mu=c(rep(0,n_noise),rep(0, n_signal)), Sigma=sigma)
  x2 <- mvrnorm(n=n, mu=c(rep(0,n_noise),rep(1, n_signal)*signal_strenght), sigma)
  x <- rbind(x1, x2)
  y <- c(rep(0, nrow(x1)), rep(1, nrow(x2)))
  df <- as.data.frame(cbind(y, x))
}
```

Putting it all together to run simulations

```{r, cache=TRUE}
gpc_comparison_mixedgaussian_generator <- function(prop){
  compare_gpc_full_vs_reduced(mixed_gaussians_data_generator,
                              n = 100,
                              prop = prop)
}

gpc_comparison_mixedgaussian_generator_null <- function(prop){
  compare_gpc_full_vs_reduced(mixed_gaussians_data_generator,
                              n = 100,
                              prop = prop,
                              shuffle.predictions = TRUE)
}

n_vec <- c(0.1, 0.25, 0.5, 0.75, 1)
n_vec <- seq(0.1, 0.9, by=0.2)

results_gpc_simulated <- calculate_proportion_signifficant(
                                 gpc_comparison_mixedgaussian_generator,
                                 parameters = n_vec,
                                 n_sims = n_sims,
                                 ncores = 4)
plot_results_lines(results_gpc_simulated)
```

```{r, cache=TRUE}
results_gpc_simulated_null <- calculate_proportion_signifficant(
                                 gpc_comparison_mixedgaussian_generator_null,
                                 parameters = seq(0.1, 1, length.out = 6),
                                 n_sims = n_sims*10,
                                 ncores = 4)
plot_results_lines(results_gpc_simulated_null)
```


### Real data - OASIS

Now we create the data generator. This generator takes the number of subjects *n* as an input and returns a random sample of size *n* with the same number of subjects for each target group (male or female). We also prepare a generator function to create the null data i.e., data with shuffled target values. 

```{r, cache=TRUE}
oasis_data_gender_gm <- load_oasis(y = 'gender', modality = 'gm')
oasis_data_gender_wm <- load_oasis(y = 'gender', modality = 'wm')
oasis_data_diag_gm <- load_oasis(y = 'diagnosis', modality = 'gm')
oasis_data_diag_wm <- load_oasis(y = 'diagnosis', modality = 'wm')


oasis_balanced_data_generator <- function(oasis_data, n, shuffle.y=FALSE){
  idx0 <- which(oasis_data$y %in% 0, arr.ind = T)
  idx1 <- which(oasis_data$y %in% 1, arr.ind = T)
  selected <- c(sample(idx0, n/2), sample(idx1, n/2))
  return_data <- oasis_data[selected,]
  if (shuffle.y){
    return_data$y <- sample(return_data$y)
  }
  return(return_data)
}


oasis_gender_gm_generator <- function(n, shuffle.y = FALSE){
  oasis_balanced_data_generator(oasis_data_gender_gm, n, shuffle.y = shuffle.y)
}
oasis_gender_wm_generator <- function(n, shuffle.y = FALSE){
  oasis_balanced_data_generator(oasis_data_gender_wm, n, shuffle.y = shuffle.y)
}
oasis_diag_gm_generator <- function(n, shuffle.y = FALSE){
  oasis_balanced_data_generator(oasis_data_diag_gm, n, shuffle.y = shuffle.y)
}
oasis_diag_wm_generator <- function(n, shuffle.y = FALSE){
  oasis_balanced_data_generator(oasis_data_diag_wm, n, shuffle.y = shuffle.y)
}


gpc_comparison_oasis_gender_gm_generator <- function(prop){
  compare_gpc_full_vs_reduced(oasis_gender_gm_generator,
                              n = 100,
                              prop = prop)
}
gpc_comparison_oasis_gender_wm_generator <- function(prop){
  compare_gpc_full_vs_reduced(oasis_gender_wm_generator,
                              n = 100,
                              prop = prop)
}
gpc_comparison_oasis_diag_gm_generator <- function(prop){
  compare_gpc_full_vs_reduced(oasis_diag_gm_generator,
                              n = 50,
                              prop = prop)
}
gpc_comparison_oasis_diag_wm_generator <- function(prop){
  compare_gpc_full_vs_reduced(oasis_diag_wm_generator,
                              n = 50,
                              prop = prop)
}


gpc_comparison_oasis_gender_gm_generator_null <- function(prop){
  compare_gpc_full_vs_reduced(oasis_gender_gm_generator,
                              n = 100,
                              prop = prop,
                              shuffle.predictions = T)
}
gpc_comparison_oasis_gender_wm_generator_null <- function(prop){
  compare_gpc_full_vs_reduced(oasis_gender_wm_generator,
                              n = 100,
                              prop = prop,
                              shuffle.predictions = T)
}
gpc_comparison_oasis_diag_gm_generator_null <- function(prop){
  compare_gpc_full_vs_reduced(oasis_diag_gm_generator,
                              n = 50,
                              prop = prop,
                              shuffle.predictions = T)
}
gpc_comparison_oasis_diag_wm_generator_null <- function(prop){
  compare_gpc_full_vs_reduced(oasis_diag_wm_generator,
                              n = 50,
                              prop = prop,
                              shuffle.predictions = T)
}


results_gpc_oasis_gender_gm <- calculate_proportion_signifficant(
                                 gpc_comparison_oasis_gender_gm_generator,
                                 parameters = n_vec,
                                 n_sims = n_sims,
                                 ncores = 4)
results_gpc_oasis_gender_wm <- calculate_proportion_signifficant(
                                 gpc_comparison_oasis_gender_wm_generator,
                                 parameters = n_vec,
                                 n_sims = n_sims,
                                 ncores = 4)
results_gpc_oasis_diag_gm <- calculate_proportion_signifficant(
                                 gpc_comparison_oasis_diag_gm_generator,
                                 parameters = n_vec,
                                 n_sims = n_sims,
                                 ncores = 4)
results_gpc_oasis_diag_wm <- calculate_proportion_signifficant(
                                 gpc_comparison_oasis_diag_wm_generator,
                                 parameters = n_vec,
                                 n_sims = n_sims,
                                 ncores = 4)


results_gpc_oasis_gender_gm_null <- calculate_proportion_signifficant(
                                 gpc_comparison_oasis_gender_gm_generator_null,
                                 parameters = n_vec,
                                 n_sims = n_sims*10,
                                 ncores = 4)
results_gpc_oasis_gender_wm_null <- calculate_proportion_signifficant(
                                 gpc_comparison_oasis_gender_wm_generator_null,
                                 parameters = n_vec,
                                 n_sims = n_sims*10,
                                 ncores = 4)
results_gpc_oasis_diag_gm_null <- calculate_proportion_signifficant(
                                 gpc_comparison_oasis_diag_gm_generator_null,
                                 parameters = n_vec,
                                 n_sims = n_sims*10,
                                 ncores = 4)
results_gpc_oasis_diag_wm_null <- calculate_proportion_signifficant(
                                 gpc_comparison_oasis_diag_wm_generator_null,
                                 parameters = n_vec,
                                 n_sims = n_sims*10,
                                 ncores = 4)


plot_results_lines(results_gpc_oasis_gender_gm)
plot_results_lines(results_gpc_oasis_gender_wm)
plot_results_lines(results_gpc_oasis_diag_gm)
plot_results_lines(results_gpc_oasis_diag_wm)

plot_results_lines(results_gpc_oasis_gender_gm_null)
plot_results_lines(results_gpc_oasis_gender_wm_null)
plot_results_lines(results_gpc_oasis_diag_gm_null)
plot_results_lines(results_gpc_oasis_diag_wm_null)

```

### Real data - abide

```{r, cache=TRUE}
abide_data <- load_abide()

abide_balanced_data_generator <- function(n, shuffle.y=FALSE){
  idx0 <- which(abide_data$y %in% 0, arr.ind = T)
  idx1 <- which(abide_data$y %in% 1, arr.ind = T)
  selected <- c(sample(idx0, n/2), sample(idx1, n/2))
  selected_data <- abide_data[selected,]
  selected_data <- selected_data[,apply(selected_data, 2, sd) != 0]
  if (shuffle.y){
    selected_data$y <- sample(selected_data$y)
  }
  return(selected_data)
}

abide_balanced_data_generator_null <- function(n){
  abide_balanced_data_generator(n, shuffle.y = TRUE)
}

gpc_comparison_abide_generator <- function(prop){
  compare_gpc_full_vs_reduced(abide_balanced_data_generator,
                              n = 100,
                              prop = prop)
}

gpc_comparison_abide_generator_null <- function(prop){
  compare_gpc_full_vs_reduced(abide_balanced_data_generator,
                              n = 100,
                              prop = prop,
                              shuffle.predictions = T)
}


results_gpc_abide <- calculate_proportion_signifficant(
                                 gpc_comparison_abide_generator,
                                 parameters = n_vec,
                                 n_sims = n_sims,
                                 ncores = 4)
results_gpc_abide_null <- calculate_proportion_signifficant(
                                 gpc_comparison_abide_generator_null,
                                 parameters = n_vec,
                                 n_sims = n_sims*10,
                                 ncores = 4)


plot_results_lines(results_gpc_abide)
plot_results_lines(results_gpc_abide_null)


```

## non imaging datasets

```{r, cache=TRUE}
library(mlbench)
data("PimaIndiansDiabetes")
names(PimaIndiansDiabetes)[9] <- 'y'
PimaIndiansDiabetes$y <- as.numeric(PimaIndiansDiabetes$y) - 1


pima_balanced_data_generator <- function(n, shuffle.y=FALSE){
  idx0 <- which(PimaIndiansDiabetes$y %in% 0, arr.ind = T)
  idx1 <- which(PimaIndiansDiabetes$y %in% 1, arr.ind = T)
  selected <- c(sample(idx0, n/2), sample(idx1, n/2))
  selected_data <- PimaIndiansDiabetes[selected,]
  selected_data <- selected_data[,apply(selected_data, 2, sd) != 0]
  if (shuffle.y){
    selected_data$y <- sample(selected_data$y)
  }
  return(selected_data)
}

pima_balanced_data_generator_null <- function(n){
  pima_balanced_data_generator(n, shuffle.y = TRUE)
}

gpc_comparison_pima_generator <- function(prop){
  compare_gpc_full_vs_reduced(pima_balanced_data_generator,
                              n = 100,
                              prop = prop)
}

gpc_comparison_pima_generator_null <- function(prop){
  compare_gpc_full_vs_reduced(pima_balanced_data_generator,
                              n = 100,
                              prop = prop,
                              shuffle.predictions = T)
}


results_gpc_pima <- calculate_proportion_signifficant(
                                 gpc_comparison_pima_generator,
                                 parameters = n_vec,
                                 n_sims = n_sims,
                                 ncores = 4)
results_gpc_pima_null <- calculate_proportion_signifficant(
                                 gpc_comparison_pima_generator_null,
                                 parameters = n_vec,
                                 n_sims = n_sims*10,
                                 ncores = 4)


plot_results_lines(results_gpc_pima)
plot_results_lines(results_gpc_pima_null)
```

```{r, cache=TRUE}
data("Sonar")
Sonar$Class <- as.numeric(Sonar$Class) -1
names(Sonar)[61] <- 'y'


sonar_balanced_data_generator <- function(n, shuffle.y=FALSE){
  idx0 <- which(Sonar$y %in% 0, arr.ind = T)
  idx1 <- which(Sonar$y %in% 1, arr.ind = T)
  selected <- c(sample(idx0, n/2), sample(idx1, n/2))
  selected_data <- Sonar[selected,]
  selected_data <- selected_data[,apply(selected_data, 2, sd) != 0]
  if (shuffle.y){
    selected_data$y <- sample(selected_data$y)
  }
  return(selected_data)
}

sonar_balanced_data_generator_null <- function(n){
  sonar_balanced_data_generator(n, shuffle.y = TRUE)
}

gpc_comparison_sonar_generator <- function(prop){
  compare_gpc_full_vs_reduced(sonar_balanced_data_generator,
                              n = 50,
                              prop = prop)
}

gpc_comparison_sonar_generator_null <- function(prop){
  compare_gpc_full_vs_reduced(sonar_balanced_data_generator,
                              n = 50,
                              prop = prop,
                              shuffle.predictions = T)
}


results_gpc_sonar <- calculate_proportion_signifficant(
                                 gpc_comparison_sonar_generator,
                                 parameters = n_vec,
                                 n_sims = n_sims,
                                 ncores = 4)
results_gpc_sonar_null <- calculate_proportion_signifficant(
                                 gpc_comparison_sonar_generator_null,
                                 parameters = n_vec,
                                 n_sims = n_sims*10,
                                 ncores = 4)


plot_results_lines(results_gpc_sonar)
plot_results_lines(results_gpc_sonar_null)
```


```{r, cache=TRUE}
library(kernlab)
data("musk")
musk$Class <- as.numeric(musk$Class) - 1
names(musk)[167] <- 'y'


musk_balanced_data_generator <- function(n, shuffle.y=FALSE){
  idx0 <- which(musk$y %in% 0, arr.ind = T)
  idx1 <- which(musk$y %in% 1, arr.ind = T)
  selected <- c(sample(idx0, n/2), sample(idx1, n/2))
  selected_data <- musk[selected,]
  selected_data <- selected_data[,apply(selected_data, 2, sd) != 0]
  if (shuffle.y){
    selected_data$y <- sample(selected_data$y)
  }
  return(selected_data)
}

musk_balanced_data_generator_null <- function(n){
  musk_balanced_data_generator(n, shuffle.y = TRUE)
}

gpc_comparison_musk_generator <- function(prop){
  compare_gpc_full_vs_reduced(musk_balanced_data_generator,
                              n = 100,
                              prop = prop)
}

gpc_comparison_musk_generator_null <- function(prop){
  compare_gpc_full_vs_reduced(musk_balanced_data_generator,
                              n = 100,
                              prop = prop,
                              shuffle.predictions = T)
}


results_gpc_musk <- calculate_proportion_signifficant(
                                 gpc_comparison_musk_generator,
                                 parameters = n_vec,
                                 n_sims = n_sims,
                                 ncores = 4)
results_gpc_musk_null <- calculate_proportion_signifficant(
                                 gpc_comparison_musk_generator_null,
                                 parameters = n_vec,
                                 n_sims = n_sims*10,
                                 ncores = 4)


plot_results_lines(results_gpc_musk)
plot_results_lines(results_gpc_musk_null)
```

```{r, cache=TRUE}
data("spam")
spam$type <- as.numeric(spam$type) - 1
names(spam)[58] <- 'y'


spam_balanced_data_generator <- function(n, shuffle.y=FALSE){
  idx0 <- which(spam$y %in% 0, arr.ind = T)
  idx1 <- which(spam$y %in% 1, arr.ind = T)
  selected <- c(sample(idx0, n/2), sample(idx1, n/2))
  selected_data <- spam[selected,]
  selected_data <- selected_data[,apply(selected_data, 2, sd) != 0]
  if (shuffle.y){
    selected_data$y <- sample(selected_data$y)
  }
  return(selected_data)
}

spam_balanced_data_generator_null <- function(n){
  spam_balanced_data_generator(n, shuffle.y = TRUE)
}

gpc_comparison_spam_generator <- function(prop){
  compare_gpc_full_vs_reduced(spam_balanced_data_generator,
                              n = 100,
                              prop = prop)
}

gpc_comparison_spam_generator_null <- function(prop){
  compare_gpc_full_vs_reduced(spam_balanced_data_generator,
                              n = 100,
                              prop = prop,
                              shuffle.predictions = T)
}


results_gpc_spam <- calculate_proportion_signifficant(
                                 gpc_comparison_spam_generator,
                                 parameters = n_vec,
                                 n_sims = n_sims,
                                 ncores = 4)
results_gpc_spam_null <- calculate_proportion_signifficant(
                                 gpc_comparison_spam_generator_null,
                                 parameters = n_vec,
                                 n_sims = n_sims*10,
                                 ncores = 4)


plot_results_lines(results_gpc_spam)
plot_results_lines(results_gpc_spam_null)
```

## plotting

```{r, fig.width=7.4, fig.height=6}
plot_results_lines <- function(alphas, id_name='parameter'){
  lines_colors <- c("black", "#375e97", "#fb6542", 'black')
  melted_alphas <- melt(alphas, id.vars = id_name)
  names(melted_alphas)[1] <- 'x'
  
  ggplot(melted_alphas, aes(x=x, y=value, color=variable, linetype=variable)) +
    geom_line()  +
    geom_point(aes(shape=variable, size=variable)) +
    scale_color_manual(values=lines_colors) +
    scale_linetype_manual(values = c('solid', 'solid','solid',  'dashed')) +
    scale_shape_manual(values=c(19,19,19,1)) +
    scale_size_manual(values=c(2,2,2,4)) +
    
    theme_minimal_grid() +
    scale_y_continuous(limits=c(0,1), breaks=seq(0, 1, 0.2))+
    scale_x_continuous(limits=c(0,1), breaks=seq(0.1, 0.9, 0.2))+
    theme(aspect.ratio = 1,
          # text=element_text(size=7),
          # axis.text = element_text(size=7),
          legend.position = 'bottom',
          legend.justification = c(0.5, 0.5),
          legend.title=element_blank(),
          legend.key.width = unit(10, 'mm')) +
    ylab('Proportion p < 0.05') +
    xlab('Relative training set size')
}


pubtheme <- theme(aspect.ratio = 1,
                  text=element_text(size=7),
                  axis.text = element_text(size=7), 
                  legend.position = 'bottom',
                  legend.justification = c(0.5, 0.5))

library(ggpubr)


p1 <- plot_results_lines(results_comp_n) + 
  pubtheme + scale_x_continuous(limits = range(results_comp_n$parameter),
                                breaks = results_comp_n$parameter) +
  xlab('n')
p2 <- plot_results_lines(results_gpc_simulated) + pubtheme
p3 <- plot_results_lines(results_gpc_oasis_gender_gm) + pubtheme
p4 <- plot_results_lines(results_gpc_oasis_gender_wm) + pubtheme
p5 <- plot_results_lines(results_gpc_oasis_diag_gm) + pubtheme
p6 <- plot_results_lines(results_gpc_oasis_diag_wm) + pubtheme
p7 <- plot_results_lines(results_gpc_abide) + pubtheme
p8 <- plot_results_lines(results_gpc_pima) + pubtheme
p9 <- plot_results_lines(results_gpc_musk) + pubtheme
p10 <- plot_results_lines(results_gpc_sonar) + pubtheme
p11 <- plot_results_lines(results_gpc_spam) + pubtheme

p_all <- ggarrange(plotlist = list(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11),
          common.legend = T, 
          labels = 'AUTO',
          font.label = list(size=8),
          legend='bottom')
p_all

# ggsave('power_comparison.tiff', p_all, width = 190, height = 150, units = 'mm', dpi=300)
# ggsave('power_comparison.tiff', p_all, width = 6, height = 4.7, units = 'in')
```


```{r, fig.width=7.5, fig.height=6}
p1 <- plot_results_lines(results_comp_n_null) + 
  pubtheme + scale_x_continuous(limits = range(results_comp_n$parameter),
                                breaks = results_comp_n$parameter) + ylim(c(0,0.1))
p2 <- plot_results_lines(results_gpc_simulated_null) + pubtheme + ylim(c(0,0.1))
p3 <- plot_results_lines(results_gpc_oasis_gender_gm_null) + pubtheme + ylim(c(0,0.1))
p4 <- plot_results_lines(results_gpc_oasis_gender_wm_null) + pubtheme + ylim(c(0,0.1))
p5 <- plot_results_lines(results_gpc_oasis_diag_gm_null) + pubtheme + ylim(c(0,0.1))
p6 <- plot_results_lines(results_gpc_oasis_diag_wm_null) + pubtheme + ylim(c(0,0.1))
p7 <- plot_results_lines(results_gpc_abide_null) + pubtheme + ylim(c(0,0.1))
p8 <- plot_results_lines(results_gpc_pima_null) + pubtheme + ylim(c(0,0.1))
p9 <- plot_results_lines(results_gpc_musk_null) + pubtheme + ylim(c(0,0.1))
p10 <- plot_results_lines(results_gpc_sonar_null) + pubtheme + ylim(c(0,0.1))
p11 <- plot_results_lines(results_gpc_spam_null) + pubtheme + ylim(c(0,0.1))

ggarrange(plotlist = list(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11),
          common.legend = T,
          legend='bottom')
```
